---
title: "Project week 2 course Reproducible Research"
author: "Jessica Quintero López"
date: "2/3/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    code_folding: hide
    theme: sandstone
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## <font color ="green">**Reading Data**</font> 

```{r}
data <- read.csv('activity.csv', sep=",")
```

## <font color ="green">**Histogram of the total number of steps taken each day**</font> 

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
steps <- na.omit(data %>% group_by(as.Date(date)) %>% summarise(sum(steps)))
colnames(steps) <- c("Date","Steps")
hist(steps$Steps, col="darkblue",border = "darkblue", main="Number of steps taken each day",
     xlab = "Steps", breaks=12)
```

##  <font color ="green">**Mean and median number of steps taken each day**</font> {.tabset}

###  <font color ="green">**Mean number of steps taken each day**</font>

```{r}
mean_steps <- mean(na.omit(steps$Steps))
mean_steps
```
###  <font color ="green">**Median number of steps taken each day**</font>

```{r}
median_steps <- median(na.omit(steps$Steps))
median_steps
```
## <font color ="green">**Time series plot of the average number of steps taken**</font> 

```{r,message=FALSE, warning=FALSE}
library(tidyverse)
s_time <- na.omit(data %>% group_by(as.Date(date)) %>% summarise(mean(steps)))
colnames(s_time) <- c("Date","MeanSteps")
```

```{r}
plot(s_time$Date,s_time$MeanSteps, type="l", col="red",
     xlab="Time", ylab="Steps", main="Time series plot of the average number of steps taken")
```

## <font color ="green">**The 5-minute interval that, on average, contains the maximum number of steps**</font> 


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
s_time <- na.omit(data %>% group_by(as.Date(date)) %>% summarise(mean(steps)))
colnames(s_time) <- c("Date","MeanSteps")
interval_row <- which.max(s_time$MeanSteps)
max_interval <- s_time[interval_row,2]
max_interval
```


## <font color ="green">**Code to describe and show a strategy for imputing missing data**</font> 

```{r, message=FALSE, warning=FALSE}
library(modeest)
data$steps[is.na(data$steps)] <- mlv(data$steps, method = "mfv")[1]
head(data)
sum(is.na(data))
```

## <font color ="green">**Histogram of the total number of steps taken each day after missing values are imputed**</font> 

```{r}
hist(data$steps, col="purple",border = "purple", main="Number of steps taken each day",
     xlab = "Steps")
```
## <font color="green> **Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends** </font>

```{r}
data$date <- as.Date(data$date)
data$weekday <- weekdays(data$date)
data$day_type <- ifelse(data$weekday=="sábado" |
                                           data$weekday=="domingo","Weekend","Weekday")
data$day_type <- factor(data$day_type)
```

```{r}
day_t<- aggregate(steps ~ interval + day_type, data=data, mean)
library(ggplot2)
ggplot(day_t, aes(interval, steps)) + 
        geom_line() + 
        facet_grid(day_type ~ .) +
        xlab("5-minute intervals") + 
        ylab("Avarage number of steps taken") +
        ggtitle("Weekdays and weekends activity patterns")
```

